### 2017-08-30 paper.tex
:art: Collapse of a dense granular column

@@

Dr.\ Krishna Kumar\\
Department of Engineering \\
University of Cambridge \\
Cambridge, CB2 1PZ\\
Tel.: +44(0) 1223 748 589\\
e-mail: \url{kks32@cam.ac.uk}\\
Web: \url{www.cb-geo.com}\\

>>>>

\section{Collapse of a dense granular column}
\Cref{fig:dense-s5-r09} shows the snapshots of the flow evolution of a dense granular column (aspect ratio 0.8), which has an initial packing density of $\Phi = 83\%$, on a \SI{5}{\degree} slope. The failure begins at the toe end of the column, and the shear-failure surface propagates into the column at an angle of about \SI{45 - 50}{\degree} The failure is due to collapse of the flank.  Force chains can be observed in the static region of the column. Once the material is destabilised, the surface of the flowing granular mass interacts with the surrounding fluid, resulting in the formation of turbulent vortices. The vortices are formed only during the horizontal acceleration phase and interact with the grains at the surface resulting in an irregular free surface. As the granular material ceases to flow, force chains restart to develop at the flow front, revealing consolidation of the granular mass with an increase in the shear resistance. The vortices formed during the collapse start to raise above the settled granular mass.

\begin{figure}[htpb]
\includegraphics[width=0.97\columnwidth]{figs/dense_s5_r09}
\caption{Evolution of a dense sand column (aspect ratio 0.8) on a slope of \SI{5}{\degree}. }
\label{fig:dense-s5-r09}
\end{figure}

@@

Prof.\ Kenichi Soga\\
Department of Engineering \\
University of Cambridge \\
Cambridge, CB2 1PZ\\
Tel.: +44(0) 1223 748 589\\
e-mail: \url{kks32@cam.ac.uk}\\
Web: \url{www.cb-geo.com}\\
